### 2025-04-09 correlation.qmd
20250408 - interactive scatterplot with best-fit line

@@

Please note that the $p$-value for a correlation assumes that the observations are independent—in particular, that the residuals are not correlated.
However, the observations are not independent in the `player_stats_seasonal` dataframe used above, because the same player has multiple rows—one row corresponding to each season they played.
This non-independence violates the traditional assumptions of the significance test of a correlation.
We could address this assumption by analyzing only one season from each player or by estimating the significance of the correlation coefficient using cluster-robust standard errors.
For simplicity, we present results above from the whole dataframe.
In @sec-mixedModels, we discuss [mixed model](#sec-mixedModels) approaches that handle repeated measures and other data that violate assumptions of non-independence that are shared by correlation and [multiple regression](#sec-multipleRegression); assumptions of [multiple regression](#sec-multipleRegression) are described in @sec-assumptionsRegression.
Below, we demonstrate how to account for non-independence of observations using cluster-robust standard errors.

```{r}

```

>>>>

### Scatterplot with Best-Fit Line

A scatterplot with best-fit line

```{r}
#| label: fig-scatterplotWithBestFitLine
#| fig-cap: "Scatterplot of Fantasy Points (Season) by Player Height With Best-Fit Line."
#| fig-alt: "Scatterplot of Fantasy Points (Season) by Player Height With Best-Fit Line."

plot_scatterplot <- ggplot2::ggplot(
  data = player_stats_seasonal %>% filter(position %in% c("QB","RB","WR","TE")),
  aes(
    x = height,
    y = fantasyPoints)) +
  geom_point(
    aes(
      text = player_display_name, # add player name for mouse over tooltip
      label = season), # add season for mouse over tooltip
    alpha = 0.05) +
  geom_smooth(
    method = "lm",
    color = "black") +
  geom_smooth() +
  coord_cartesian(
    ylim = c(0,NA),
    expand = FALSE) +
  labs(
    x = "Player Height (Inches)",
    y = "Fantasy Points (Season)",
    title = "Fantasy Points (Season) by Player Height",
    subtitle = "(Among QBs, RBs, WRs, and TEs)"
  ) +
  theme_classic()

plotly::ggplotly(plot_scatterplot)
```

```{r}
#| label: fig-scatterplotByPositionWithBestFitLine
#| fig-cap: "Scatterplot of Fantasy Points (Season) by Player Height and Position With Best-Fit Line."
#| fig-alt: "Scatterplot of Fantasy Points (Season) by Player Height and Position With Best-Fit Line."

plot_scatterplotByPosition <- ggplot2::ggplot(
  data = player_stats_seasonal %>% filter(position %in% c("QB","RB","WR","TE")),
  aes(
    x = height,
    y = fantasyPoints,
    color = position,
    fill = position)) +
  geom_point(
    aes(
      text = player_display_name, # add player name for mouse over tooltip
      label = season), # add season for mouse over tooltip
    alpha = 0.7) +
  geom_smooth(
    method = "lm",
    color = "black") +
  geom_smooth(
    method = "loess",
    span = 0.5) +
  coord_cartesian(
    ylim = c(0,NA),
    expand = FALSE) +
  labs(
    x = "Player Height (Inches)",
    y = "Fantasy Points (Season)",
    title = "Fantasy Points (Season) by Player Height and Position"
  ) +
  theme_classic()

plotly::ggplotly(plot_scatterplotByPosition)
```

@@

Please note that the $p$-value for a correlation assumes that the observations are independent—in particular, that the residuals are not correlated.
However, the observations are not independent in the `player_stats_seasonal` dataframe used above, because the same player has multiple rows—one row corresponding to each season they played.
This non-independence violates the traditional assumptions of the significance test of a correlation.
We could address this assumption by analyzing only one season from each player or by estimating the significance of the correlation coefficient using cluster-robust standard errors.
For simplicity, we present results above from the whole dataframe.
In @sec-mixedModels, we discuss [mixed model](#sec-mixedModels) approaches that handle repeated measures and other data that violate assumptions of non-independence that are shared by correlation and [multiple regression](#sec-multipleRegression); assumptions of [multiple regression](#sec-multipleRegression) are described in @sec-assumptionsRegression.
In section @sec-multipleRegressionNonIndependence, we demonstrate how to account for non-independence of observations using cluster-robust standard errors.
